\section*{Acknowledgments}

First of all, I would like to express my gratitude to Prof. Rolf Backofen. Having the opportunity to complete my thesis in the Bioinformatics Group at the University of Freiburg is a huge step for me.  Moreover, it's an honor to have him as my examiner.

I'm also grateful to Prof. Dr. Wolfgang R. Hess for considering the role of my second examiner. His expertise and willingness to participate in this process are greatly appreciated.

My supervisors Bérénice Batut and Teresa Müller have been anchors in my research journey. Looking back to my beginnings and to the present, their influence on me at every step of the way is evident; Bérénice's insights often illuminated the direction of my research, while Teresa's pragmatism served as a compass when I felt lost. The patience they showed, especially when I was skeptical, was commendable. And the countless hours of discussion? They were the building blocks of my confidence. Their trust in me, even when I was hesitant, pushed me to aim higher and achieve more. I could not have asked for better guides on this journey than them.

I appreciate my dear friends who have been there for me throughout this process. A heartfelt shoutout to Polina – introducing me to the opportunities at the Bioinformatics Group was a turning point. I'm ever grateful for that.

The support of my family has always surrounded me, even from a distance. My parents and sister, your strong belief in me pushes me forward. Not only do you applaud my academic progress, but you are always concerned about my mental health and remind me of the importance of balance. For all this and more, I am grateful beyond words.

Thank you all for being a part of this incredible journey.